Phenotypic evolution is less constrained in complex food webs

### 2018-09-21 temporary_morgue/Manuscript/Manuscript.Rmd
getting organized

---
title: Phenotypic evolution is less constrained in complex food webs or Food-web complexity alters the fitness landscape of an insect herbivore 
author:
  - name: Matthew A. Barbour
    email: matthew.barbour@ieu.uzh.ch
    affiliation: a,b
    footnote: Corresponding Author
  - name: Christopher J. Greyson-Gaito
    affiliation: a,c
  - name: Arezoo Sootodeh
    affiliation: a
  - name: Brendan Locke
    affiliation: d
  - name: Jordi Bascompte
    affiliation: b
address:
  - code: a
    address: University of British Columbia, Department of Zoology, 6270 University Blvd., Vancouver, BC, V6T 1Z4, Canada
  - code: b
    address: University of Zurich, Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse 190, Zurich, 8057, Switzerland
  - code: c
    address: University of Guelph, Department of Integrative Biology, 50 Stone Rd. East, Guelph, ONT, N1G 2W1, Canada
  - code: d
    address: Humboldt State University, Department of Biological Sciences, 1 Harpst St., Arcata, CA, 95521, USA
abstract: "Species-interaction networks provide a mechanistic link between community and population ecology, as they describe who interacts with whom in a community. Similarly, fitness landscapes provide a mechanistic link between population and evolutionary ecology, as they describe the influence of natural selection on phenotypic variation. It remains unclear, however, how the network structure of species interactions shapes the fitness landscape. Understanding this relationship is likely key for developing a predictive ecology across scales of biological organization. To examine the relationship between network structure and the fitness landscape, we conducted a field experiment that manipulated the network of trophic interactions (simple vs. complex) influencing the fitness of an insect herbivore. We then quantified the fitness landscape of herbivores in each treatment by measuring herbivore survival as a function of multiple phenotypic traits. We found that more traits were under selection in the simple vs. complex trophic network. This occurred because different natural enemies impose different selection pressures on herbivore traits, thereby minimizing relative fitness differences among herbivore phenotypes in the complex network. Our work suggests that more complex trophic networks allow phenotypic variation to persist, which could facilitate subsequent adaptive evolution of populations to environmental change."

bibliography: mybibfile.bib
output: rticles::elsevier_article
---

@@

## Number of bootstrap replicates
n_boots_analysis <- 30 #1000
n_boots_plots <- 100

@@

To characterize the net effects of food-web complexity on the slope and curvature of *Iteomyia*'s fitness landscape, we took advantage of existing theory that links selection surfaces of individuals to the fitness landscape of the population (Phillips & Arnold 1998, Arnold 2003). Specifically, the slope of the fitness landscape corresponds to the column vector of directional selection gradients:
$$\text{Slope} = \beta = \begin{pmatrix} \beta_{diam} \\ \beta_{clutch} \\ \beta_{pref} \end{pmatrix} $$
whereas the curvature of the fitness landscape is a function of the directional, nonlinear, and correlational selection gradients: 
$$\text{Curvature} = \gamma - \beta \beta^T$$
where \gamma represents the matrix of nonlinear and correlational selection gradients:
$$\gamma = \begin{pmatrix} \gamma_{diam,diam}&& \\ \gamma_{diam,clutch}&\gamma_{clutch,clutch}& \\ \gamma_{diam,pref} & \gamma_{clutch,pref} &\gamma_{pref,pref} \end{pmatrix}$$
Note that we ommitted the upper triangle of the matrix for clarity since it is simply the reflection of the lower triangle. Assuming that there is additive genetic variance and covariance between these traits under selection, then the slope and curvature of the fitness landscape give insight to how the population's mean trait value will change in the next generation as well as how additive genetic variance and covariance changes within a generation.

@@

# There is clear evidence that the traits we have measured influence larva survival
anova(null_glmer, trait_glmer, test="Chisq")
rsquared(list(null_glmer, trait_glmer))

@@

```{r Get Selection Gradients for GLMER}

@@

```{r Get Selection Gradients for GLMER}

## Estimate mean fitness and mean brackets (ref: Janzen and Stern 1998)
complex_predict <- predict(foodweb_glmer, newdata=filter(gall_selection.df, Foodweb=="Complex"), type="response")
complex_mean_brackets <- mean(complex_predict * (1 - complex_predict))
complex_mean_fitness <- mean(complex_predict)

@@

# test whether using alternative models influences estimes of mean "brackets" and mean fitness 
alt_predict <- predict(foodweb_glmer_2, newdata=filter(gall_selection.df, Foodweb=="Complex"), type="response")
alt_mean_brackets <- mean(alt_predict * (1 - alt_predict))
alt_mean_fitness <- mean(alt_predict)
complex_mean_brackets - alt_mean_brackets
complex_mean_fitness - alt_mean_fitness

# Altering so that coefficients that differed between food webs are reported, but ones that weren't are simply reported as there baselines.
# Added multiplier column so that quadratic (nonlinear) selection coefficients can be multiplied by 2 for fair comparison
foodweb_tidy_4 <- tidy(foodweb_glmer_4, conf.int=T) %>%
  filter(term %in% c("(Intercept)","FoodwebSimple","sc.Diam","sc.log1p.Pref")) %>%
  mutate(Foodweb = c("Complex","Simple","Common","Common"),
         Selection_form = c(NA,NA,"Directional","Directional"),
         Multiplier = c(1,1,1,1))

foodweb_tidy_3 <- tidy(foodweb_glmer_3, conf.int=T) %>%
  filter(term %in% c("sc.log.Clutch","sc.Diam:sc.log.Clutch","sc.Diam:sc.log1p.Pref","sc.log.Clutch:sc.log1p.Pref","FoodwebSimple:sc.log.Clutch")) %>%
  mutate(Foodweb = c("Complex","Common","Common","Common","Simple"),
         Selection_form = c("Directional","Correlational","Correlational","Correlational","Directional"),
         Multiplier = c(1,1,1,1,1))

foodweb_tidy_2 <- tidy(foodweb_glmer_2, conf.int=T) %>%
  filter(term %in% c("I(sc.Diam^2)","I(sc.log.Clutch^2)","sc.Diam:sc.log.Clutch:sc.log1p.Pref")) %>%
  mutate(Foodweb = c("Common","Common","Common"),
         Selection_form = c("Nonlinear","Nonlinear","Correlational"),
         Multiplier = c(2,2,1))

foodweb_tidy <- tidy(foodweb_glmer, conf.int=T) %>%
  filter(term %in% c("I(sc.log1p.Pref^2)","FoodwebSimple:I(sc.log1p.Pref^2)")) %>%
  mutate(Foodweb = c("Complex","Simple"),
         Selection_form = c("Nonlinear","Nonlinear"),
         Multiplier = c(2,2))

get_alphas <- bind_rows(foodweb_tidy_4, foodweb_tidy_3, foodweb_tidy_2, foodweb_tidy) %>% 
  mutate(alpha = round(estimate, 3), alpha_2.5 = round(conf.low, 3), alpha_97.5 = round(conf.high, 3), P = round(p.value, 3)) %>%
  select(term, Foodweb, alpha, alpha_2.5, alpha_97.5, P, Selection_form, Multiplier)

complex_alphas <- filter(get_alphas, Foodweb == "Complex"); complex_alphas
simple_alphas <- filter(get_alphas, Foodweb == "Simple"); simple_alphas
common_alphas <- filter(get_alphas, Foodweb == "Common"); common_alphas

## NEED TO DUPLICATE COMMON COEFFICIENTS, AS THIS WILL MAKE IT EASIEST TO KEEP THE SAME CODE AND CALCULATE SELECTION GRADIENTS

# remember that the selection gradients need to be multiplied by 2 for quadratic gradients.
coefs_gradients.complex_simple <- data.frame(Variable = complex_alphas$term, 
           a_complex = complex_alphas$alpha,
           a_complex_2.5 = complex_alphas$alpha_2.5,
           a_complex_97.5 = complex_alphas$alpha_97.5,
           a_simple = complex_alphas$alpha + simple_alphas$alpha,        # the baseline is the coefficient in the complex food web, therefore we simply add the corresponding coefficients from the simple food web to get the coefficients on the same scale as the complex web
           a_simple_2.5 = complex_alphas$alpha + simple_alphas$alpha_2.5, # same addition process for confidence intervals
           a_simple_97.5 = complex_alphas$alpha + simple_alphas$alpha_97.5,
           Selection_form = complex_alphas$Selection_form,                # doesn't matter whether simple_ or complex_alphas is chosen for Selection_form or Multiplier
           Multiplier = complex_alphas$Multiplier) %>%                     
  mutate(b_complex = round(complex_mean_brackets * a_complex / complex_mean_fitness, 3) * Multiplier,
         b_complex_2.5 = round(complex_mean_brackets * a_complex_2.5 / complex_mean_fitness, 3) * Multiplier,
         b_complex_97.5 = round(complex_mean_brackets * a_complex_97.5 / complex_mean_fitness, 3) * Multiplier,
         b_simple = round(simple_mean_brackets * a_simple / simple_mean_fitness, 3) * Multiplier,
         b_simple_2.5 = round(simple_mean_brackets * a_simple_2.5 / simple_mean_fitness, 3) * Multiplier,
         b_simple_97.5 = round(simple_mean_brackets * a_simple_97.5 / simple_mean_fitness, 3) * Multiplier)

# duplicate common alpha coefficients for complex and simple food webs
coefs_gradients.common <- data.frame(Variable = common_alphas$term, 
           a_complex = common_alphas$alpha,
           a_complex_2.5 = common_alphas$alpha_2.5,
           a_complex_97.5 = common_alphas$alpha_97.5,
           a_simple = common_alphas$alpha,        
           a_simple_2.5 = common_alphas$alpha_2.5, 
           a_simple_97.5 = common_alphas$alpha_97.5,
           Selection_form = common_alphas$Selection_form,
           Multiplier = common_alphas$Multiplier) %>% 
  # for the selection gradients though, calculate them on the scale of relative fitness for each food-web treatment
  mutate(b_complex = round(complex_mean_brackets * a_complex / complex_mean_fitness, 3) * Multiplier,
         b_complex_2.5 = round(complex_mean_brackets * a_complex_2.5 / complex_mean_fitness, 3) * Multiplier,
         b_complex_97.5 = round(complex_mean_brackets * a_complex_97.5 / complex_mean_fitness, 3) * Multiplier,
         b_simple = round(simple_mean_brackets * a_simple / simple_mean_fitness, 3) * Multiplier,
         b_simple_2.5 = round(simple_mean_brackets * a_simple_2.5 / simple_mean_fitness, 3) * Multiplier,
         b_simple_97.5 = round(simple_mean_brackets * a_simple_97.5 / simple_mean_fitness, 3) * Multiplier)

coefs_gradients <- bind_rows(coefs_gradients.complex_simple, coefs_gradients.common)

bind_rows(mutate(select(coefs_gradients, Variable, a=a_complex, a_2.5=a_complex_2.5, a_97.5=a_complex_97.5), Foodweb = "Complex"),
          mutate(select(coefs_gradients, Variable, a=a_simple, a_2.5=a_simple_2.5, a_97.5=a_simple_97.5), Foodweb = "Simple")) %>%
  filter(Variable != "(Intercept)") %>%
  ggplot(., aes(x = Variable, fill=Foodweb, color=Foodweb)) +
  geom_point(aes(y=a), position = position_dodge(width=0.5)) +
  geom_errorbar(aes(ymax = a_97.5, ymin = a_2.5), position = position_dodge(width=0.5)) +
  coord_flip() +
  geom_hline(yintercept=0, linetype="dotted") +
  ylab("Regression coefficient")

@@

```{r Code Chunk to demonstrate use of latex2exp package, include=FALSE}
p<-ggplot(data=x, aes(x=y, y=x)) +
geom_errorbar(aes(ymin=x-ci, ymax=x+ci)) +
scale_x_discrete(breaks=c("label1","label2"),
               labels = lapply(sprintf('$\\alpha^\\beta$'), TeX)
p
```

What's the rationale for this test? First, we focus on galls where there is evidence of both survival and parasitism within the gall. The idea is that galls from the same clutch should be similar in size, therefore, any selection on diameter would be apparent. If we assume that this is all due to the effect of parasitism on larval development, truncating gall diameter, then we can consider this new selection gradient to be the result of confounding factors. Note that this is likely overestimates the effect of the confounding factor, since we are assuming that any heterogeneity in gall size (for this dataset) is due to a confounding effect of parasitism on gall diameter. Then, we can just subtract this value from the observed covariance between gall diameter and larval survival to adjust for the confounding effects of parasitism. This analysis should be relegated to the supplement.

Would we expect any confounding effects on nonlinear selection?
```{r Confounding effects of parasitism on gall size}
## paired t-test approach. Weakness is that it only calculates the selection differential rather than the gradient, which is more meaningful for multivariate trait evolution
pair_df <- gall_selection.df %>%
  group_by(Foodweb, Gall_Number) %>%
  mutate(mean_survival = mean(gall_survival)) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  select(Foodweb, Gall_Number, mean_survival, gall_survival, sc.Diam) %>%
  # calculate mean phenotype at gall level (before selection)
  mutate(all_mean_sc.Diam = mean(sc.Diam)) %>%
  # then filter surviving galls and calculate the mean phenotype (after selection)
  filter(gall_survival == 1) %>%
  mutate(survive_mean_sc.Diam = mean(sc.Diam)) %>%
  # only retain distinct data points and calculate the difference (selection differential)
  distinct(Foodweb, Gall_Number, all_mean_sc.Diam, survive_mean_sc.Diam) %>%
  mutate(diff_sc.Diam = survive_mean_sc.Diam - all_mean_sc.Diam)
t.test(diff_sc.Diam ~ Foodweb, pair_df)
t.test(pair_df$survive_mean_sc.Diam, pair_df$all_mean_sc.Diam, paired = T)

complex_pair_df <- gall_selection.df %>%
  filter(Foodweb=="Complex") %>%
  group_by(Gall_Number) %>%
  mutate(mean_survival = mean(gall_survival)) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  select(Gall_Number, mean_survival, gall_survival, sc.Diam) %>%
  # calculate mean phenotype at gall level (before selection)
  mutate(all_mean_sc.Diam = mean(sc.Diam)) %>%
  # then filter surviving galls and calculate the mean phenotype (after selection)
  filter(gall_survival == 1) %>%
  mutate(survive_mean_sc.Diam = mean(sc.Diam)) %>%
  # only retain distinct data points and calculate the difference (selection differential)
  distinct(Gall_Number, all_mean_sc.Diam, survive_mean_sc.Diam) %>%
  mutate(diff_sc.Diam = survive_mean_sc.Diam - all_mean_sc.Diam)
t.test(complex_pair_df$survive_mean_sc.Diam, complex_pair_df$all_mean_sc.Diam, paired = T)

ectos_pair_df <- gall_selection.df %>%
  filter(Foodweb=="Complex" & platy < 1) %>%
  group_by(Gall_Number) %>%
  mutate(mean_survival = mean(gall_survival)) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  select(Gall_Number, mean_survival, gall_survival, sc.Diam) %>%
  # calculate mean phenotype at gall level (before selection)
  mutate(all_mean_sc.Diam = mean(sc.Diam)) %>%
  # then filter surviving galls and calculate the mean phenotype (after selection)
  filter(gall_survival == 1) %>%
  mutate(survive_mean_sc.Diam = mean(sc.Diam)) %>%
  # only retain distinct data points and calculate the difference (selection differential)
  distinct(Gall_Number, all_mean_sc.Diam, survive_mean_sc.Diam) %>%
  mutate(diff_sc.Diam = survive_mean_sc.Diam - all_mean_sc.Diam)
t.test(ectos_pair_df$survive_mean_sc.Diam, ectos_pair_df$all_mean_sc.Diam, paired = T)

simple_pair_df <- gall_selection.df %>%
  filter(Foodweb=="Simple") %>%
  group_by(Gall_Number) %>%
  mutate(mean_survival = mean(gall_survival)) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  select(Gall_Number, mean_survival, gall_survival, sc.Diam) %>%
  # calculate mean phenotype at gall level (before selection)
  mutate(all_mean_sc.Diam = mean(sc.Diam)) %>%
  # then filter surviving galls and calculate the mean phenotype (after selection)
  filter(gall_survival == 1) %>%
  mutate(survive_mean_sc.Diam = mean(sc.Diam)) %>%
  # only retain distinct data points and calculate the difference (selection differential)
  distinct(Gall_Number, all_mean_sc.Diam, survive_mean_sc.Diam) %>%
  mutate(diff_sc.Diam = survive_mean_sc.Diam - all_mean_sc.Diam)
t.test(simple_pair_df$survive_mean_sc.Diam, simple_pair_df$all_mean_sc.Diam, paired = T)

## selection gradient approach
confounding_df <- gall_selection.df %>%
  group_by(Gall_Number) %>%
  mutate(mean_survival = mean(gall_survival)) %>%
  filter(mean_survival > 0, mean_survival < 1)

confounding_glmer <- glmer(gall_survival ~ Foodweb*(sc.Diam*sc.log.Clutch*sc.log1p.Pref +
                         I(sc.Diam^2) + I(sc.log.Clutch^2) + I(sc.log1p.Pref^2)) +
                         (1|Genotype/Plant_Position/Gall_Number),
                       data = confounding_df,
                       family = binomial(link = logit), control=glmerControl(optimizer = "bobyqa"))
summary(confounding_glmer)

## Replicate GLMM tests but with this confounding factor dataset ----

# No effects
drop1(confounding_glmer, test="Chisq")
summary(confounding_glmer)

# No effects on nonlinear selection acting on gall diameter, but also there is not much evidence for any effects.
confounding_glmer_2 <- update(confounding_glmer, .~. -Foodweb:I(sc.Diam^2) -Foodweb:I(sc.log.Clutch^2) -Foodweb:I(sc.log1p.Pref^2) -Foodweb:sc.Diam:sc.log.Clutch:sc.log1p.Pref)
drop1(confounding_glmer_2, test="Chisq") 
summary(confounding_glmer_2)

# Evidence of food-web treatment altering directional selection on clutch size; marginal evidence of correlational selection acting on gall size and clutch size.
confounding_glmer_3 <- update(confounding_glmer_2, .~. -I(sc.Diam^2) -I(sc.log.Clutch^2) -I(sc.log1p.Pref^2) -sc.Diam:sc.log.Clutch:sc.log1p.Pref -Foodweb:sc.Diam:sc.log.Clutch -Foodweb:sc.Diam:sc.log1p.Pref -Foodweb:sc.log.Clutch:sc.log1p.Pref)
drop1(confounding_glmer_3, test="Chisq") 
summary(confounding_glmer_3)

# Evidence that food-web treatment alters absolute fitness; evidence of directional selection acting on gall size; evidence of directional selection acting on female preference
confounding_glmer_4 <- update(confounding_glmer_3, .~. -sc.Diam:sc.log.Clutch -sc.Diam:sc.log1p.Pref -sc.log.Clutch:sc.log1p.Pref -Foodweb:sc.Diam -Foodweb:sc.log.Clutch -Foodweb:sc.log1p.Pref)
drop1(confounding_glmer_4, test="Chisq") 
summary(confounding_glmer_4)

fixef(foodweb_glmer_4)["sc.Diam"] - fixef(confounding_glmer_4)["sc.Diam"]

confounding_predict <- predict(foodweb_glmer, type="response") # not subsetting the data to complex or simple food webs, since we didn't find a difference in this subset
confounding_mean_brackets <- mean(confounding_predict * (1 - confounding_predict))
confounding_mean_fitness <- mean(confounding_predict)

# alternatively, this suggest that we are overestimating the selection gradient by this much...
round(confounding_mean_brackets * fixef(confounding_glmer_4)["sc.Diam"] / confounding_mean_fitness, 3)

diam_glmer <- glmer(gall_survival ~ Foodweb + sc.Diam +
                         (1|Genotype/Plant_Position/Gall_Number),
                       data = gall_selection.df,
                       family = binomial(link = logit), control=glmerControl(optimizer = "bobyqa"))
# note that the selection gradient and differential are virtually the same for gall diameter, given the similarity in their alpha coefficients.
fixef(diam_glmer)["sc.Diam"]
fixef(foodweb_glmer_4)["sc.Diam"]

# selection differentials
round(complex_mean_brackets * fixef(diam_glmer)["sc.Diam"] / complex_mean_fitness, 3)
round(simple_mean_brackets * fixef(diam_glmer)["sc.Diam"] / simple_mean_fitness, 3)
```

Appears to be selection for Platygaster to attack larva at high densities in more complex food webs (and a decrease in the variance). This may be a result of the poorer searching ability of parasitoids at very high gall densities, potentially due to saturation. This suggests that the fitness landscape may be dynamic because these selective effects (assuming there is heritable variation in the egg parasitoid) will change the following year. This could be an important discussion point to take into account for future work. This is different from simply altering the strength of selection, which will occur as species move along the fitness landscape, but this result suggests that the nature of the fitness landscape may actually be changing. For example, larval parasitoids could be driving selection on Platygaster to be very efficient at foraging, which will dampen once this pressure is removed, and then could also dampen the selection acting on the galling herbivore.
```{r Selection on Platygaster Traits}
platy_selection_df <- gall_selection.df %>%
  filter(egg_parasitoid > 0) 
platy_selection_df$Foodweb <- relevel(factor(platy_selection_df$Foodweb), ref = "Simple") # this represents the baseline or absence of selection by larval parasitoids

platy_Diam <- lmer(sc.Diam ~ Foodweb + (1|Genotype/Plant_Position/Gall_Number), platy_selection_df)
summary(platy_Diam)
Anova(platy_Diam, test.statistic = "F")

platy_Clutch <- lmer(sc.log.Clutch ~ Foodweb + (1|Genotype/Plant_Position), platy_selection_df)
summary(platy_Clutch)
Anova(platy_Clutch, test.statistic = "F")

platy_Pref <- lmer(sc.log1p.Pref ~ Foodweb + (1|Genotype), platy_selection_df)
summary(platy_Pref)
Anova(platy_Pref, test.statistic = "F")

## Test whether food-web complexity influences nonlinear selection with Levene's test for homogeneity of variance between groups
# Note that this result should be interepreted cautiously since we cannot account for non-independence due to plant genotype, plant ID, or gall ID
leveneTest(sc.Diam ~ Foodweb, platy_selection_df)
leveneTest(sc.log.Clutch ~ Foodweb, platy_selection_df)
leveneTest(sc.log1p.Pref ~ Foodweb, platy_selection_df)

ggplot(platy_selection_df, aes(x=Foodweb, y=sc.Diam)) +
  stat_summary(fun="mean_cl_boot", geom="errorbar", color="red", width=0.1) +
  stat_summary(fun.y="mean", geom="point", color="black", size=3)

ggplot(platy_selection_df, aes(x=Foodweb, y=sc.log.Clutch)) +
  stat_summary(fun="mean_cl_boot", geom="errorbar", color="red", width=0.1) +
  stat_summary(fun.y="mean", geom="point", color="black", size=3)

ggplot(platy_selection_df, aes(x=Foodweb, y=sc.log1p.Pref)) +
  stat_summary(fun="mean_cl_boot", geom="errorbar", color="red", width=0.1) +
  stat_summary(fun.y="mean", geom="point", color="black", size=3)
```

## Table for Alpha regression coefficients
\begin{table}[]
\begin{tabular}{lcc}
                                                                           \\ \hline
                       & \multicolumn{2}{c}{Food web (Estimate [95\% CI])} \\
\textbf{Coefficient}   & \textbf{Complex} & \textbf{Simple}                \\ \hline
$Diam$                 & `r ` [`r `-`r `] & `r ` [`r `-`r `]  \\
$Clutch$               & `r ` [`r `-`r `] & `r ` [`r `-`r `]  \\
$Pref$                 & `r ` [`r `-`r `] & `r ` [`r `-`r `]  \\
$Diam^2$               & `r ` [`r `-`r `] & `r ` [`r `-`r `]  \\
$Clutch^2$             & `r ` [`r `-`r `] & `r ` [`r `-`r `]  \\
$Pref^2$               & `r ` [`r `-`r `] & `r ` [`r `-`r `]  \\
$Diam:Clutch$          & `r ` [`r `-`r `] & `r ` [`r `-`r `]  \\
$Diam:Pref$            & `r ` [`r `-`r `] & `r ` [`r `-`r `]  \\
$Clutch:Pref$          & `r ` [`r `-`r `] & `r ` [`r `-`r `]              \\ \hline
\end{tabular}
\end{table}

## Table for Iteomyia selection gradients
\begin{table}[]
\begin{tabular}{lcc}
                                                                                  \\ \hline
                                  & \multicolumn{2}{c}{Food web (Estimate [95\% CI])} \\
\textbf{Selection gradient}       & \textbf{Complex} & \textbf{Simple}            \\ \hline
$\beta_{Diam}$                    & `r ` [`r `-`r `] & `r ` [`r `-`r `]  \\
$\beta_{Clutch}$                  & `r ` [`r `-`r `] & `r ` [`r `-`r `]  \\
$\beta_{Pref}$                    & `r ` [`r `-`r `] & `r ` [`r `-`r `]  \\
$\gamma_{Diam,Diam}$              & `r ` [`r `-`r `] & `r ` [`r `-`r `]  \\
$\gamma_{Clutch,Clutch}$          & `r ` [`r `-`r `] & `r ` [`r `-`r `]  \\
$\gamma_{Pref,Pref}$              & `r ` [`r `-`r `] & `r ` [`r `-`r `]  \\
$\gamma_{Diam,Clutch}$            & `r ` [`r `-`r `] & `r ` [`r `-`r `]  \\
$\gamma_{Diam,Pref}$              & `r ` [`r `-`r `] & `r ` [`r `-`r `]  \\
$\gamma_{Clutch,Pref}$            & `r ` [`r `-`r `] & `r ` [`r `-`r `]           \\ \hline
\end{tabular}
\end{table}

## Table for bias
\begin{table}[]
\begin{tabular}{lc}
                                                                  \\ \hline
\textbf{Type}                 & \textbf{Bias} \boldmath$s_{Diam}$ \\ \hline
Complex (All)                 & `r ` [`r `-`r `]  \\
Complex (Larval guild)        & `r ` [`r `-`r `]  \\
Simple (\textit{Platygaster}) & `r ` [`r `-`r `]                  \\ \hline
\end{tabular}
\end{table}

## Table for Platygaster selection
\begin{table}[]
\begin{tabular}{lc}
                                                                  \\ \hline
\textbf{Selection differential} & \textbf{Estimate [95\% CI]}     \\ \hline
$s_{Diam}$                      & `r ` [`r `-`r `]  \\
$s_{Clutch}$                    & `r ` [`r `-`r `]  \\
$s_{Pref}$                      & `r ` [`r `-`r `]  \\ 
$s_{Diam,Diam}$                 & `r ` [`r `-`r `]  \\
$s_{Clutch,Clutch}$             & `r ` [`r `-`r `]  \\
$s_{Pref,Pref}$                 & `r ` [`r `-`r `]                \\ \hline
\end{tabular}
\end{table}

```{r DUPLICATE for bootstrap test}

## Estimate mean fitness and mean brackets (ref: Janzen and Stern 1998)
complex_predict <- predict(foodweb_glmer, newdata=filter(gall_selection.df, Foodweb=="Complex"), type="response")
complex_mean_brackets <- mean(complex_predict * (1 - complex_predict))
complex_mean_fitness <- mean(complex_predict)

# test whether using alternative models influences estimes of mean "brackets" and mean fitness 
alt_predict <- predict(foodweb_glmer_2, newdata=filter(gall_selection.df, Foodweb=="Complex"), type="response")
alt_mean_brackets <- mean(alt_predict * (1 - alt_predict))
alt_mean_fitness <- mean(alt_predict)
complex_mean_brackets - alt_mean_brackets
complex_mean_fitness - alt_mean_fitness

>>>>

---
title: Phenotypic evolution is less constrained in complex food webs
author:
  - name: Matthew A. Barbour
    email: matthew.barbour@ieu.uzh.ch
    affiliation: a,b
    footnote: Corresponding Author
  - name: Christopher J. Greyson-Gaito
    affiliation: a,c
  - name: Arezoo Sootodeh
    affiliation: a
  - name: Brendan Locke
    affiliation: d
  - name: Jordi Bascompte
    affiliation: b
address:
  - code: a
    address: University of British Columbia, Department of Zoology, 6270 University Blvd., Vancouver, BC, V6T 1Z4, Canada
  - code: b
    address: University of Zurich, Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse 190, Zurich, 8057, Switzerland
  - code: c
    address: University of Guelph, Department of Integrative Biology, 50 Stone Rd. East, Guelph, ONT, N1G 2W1, Canada
  - code: d
    address: Humboldt State University, Department of Biological Sciences, 1 Harpst St., Arcata, CA, 95521, USA
abstract: |
  This is the abstract.

  It consists of two paragraphs.

journal: "An awesome journal"
date: "`r Sys.Date()`"
bibliography: mybibfile.bib
output: rticles::elsevier_article
---

@@

## Number of bootstrap replicates
n_boots_analysis <- 30
n_boots_plots <- 100

@@

To characterize the net effects of food-web complexity on the slope and curvature of *Iteomyia*'s fitness landscape, we took advantage of existing theory that links selection surfaces of individuals to the fitness landscape of the population (Phillips & Arnold 1998, Arnold 2003). Specifically, the slope of the fitness landscape corresponds to the column vector of directional selection gradients:

```{r LATEX NEED TO CHECK two, eval=FALSE, include=FALSE}
$$\text{Slope} = \beta = \begin{pmatrix} \beta_{diam} \\ \beta_{clutch} \\ \beta_{pref} \end{pmatrix} $$
whereas the curvature of the fitness landscape is a function of the directional, nonlinear, and correlational selection gradients: 
$$\text{Curvature} = \gamma - \beta \beta^T$$
where \gamma represents the matrix of nonlinear and correlational selection gradients:
$$\gamma = \begin{pmatrix} \gamma_{diam,diam}&& \\ \gamma_{diam,clutch}&\gamma_{clutch,clutch}& \\ \gamma_{diam,pref} & \gamma_{clutch,pref} &\gamma_{pref,pref} \end{pmatrix}$$
```

Note that we ommitted the upper triangle of the matrix for clarity since it is simply the reflection of the lower triangle. Assuming that there is additive genetic variance and covariance between these traits under selection, then the slope and curvature of the fitness landscape give insight to how the population's mean trait value will change in the next generation as well as how additive genetic variance and covariance changes within a generation.

@@

# There is clear evidence that the traits we have measured influence larva survival
anova(null_glmer, trait_glmer, test="Chisq")

@@

```{r Get Selection Gradients for GLMER, include=FALSE}

@@

```{r Parametric Bootstrapping of Regression Coefficients and Selection Gradients, cache=TRUE, include=FALSE}

## DUPLICATE ABOVE EXCEPT MODIFY FOR BOOTMER

@@

# Tidy data for beta terms
foodweb_tidy_betas <- tidy(bootMer(betas_glmer, FUN = fixef, nsim=n_boots_analysis, parallel="multicore", ncpus=32), conf.int=T) %>%
  filter(term %in% c("(Intercept)","FoodwebSimple","sc.Diam","sc.log.Clutch","sc.log1p.Pref",
                     "FoodwebSimple:sc.Diam","FoodwebSimple:sc.log.Clutch","FoodwebSimple:sc.log1p.Pref")) %>%
  mutate(Foodweb = c("Complex","Simple",rep("Complex",3),rep("Simple",3)),
         Selection_form = c(NA,NA,rep("Directional",6)),
         Multiplier = rep(1,8))

# Tidy data for gamma terms. Note that we multiply quadratic (nonlinear) selection coefficients by 2 so they are on the same scale as other coefficients (detailed explanation in Stinchcombe et al. 2008)
foodweb_tidy_gammas <- tidy(bootMer(foodweb_glmer, FUN = fixef, nsim=n_boots_analysis, parallel="multicore", ncpus=32), conf.int=T) %>%
  filter(term %in% c("I(sc.Diam^2)","I(sc.log.Clutch^2)","I(sc.log1p.Pref^2)",
                     "sc.Diam:sc.log.Clutch","sc.Diam:sc.log1p.Pref","sc.log.Clutch:sc.log1p.Pref",
                     "FoodwebSimple:I(sc.Diam^2)","FoodwebSimple:I(sc.log.Clutch^2)","FoodwebSimple:I(sc.log1p.Pref^2)",
                     "FoodwebSimple:sc.Diam:sc.log.Clutch","FoodwebSimple:sc.Diam:sc.log1p.Pref","FoodwebSimple:sc.log.Clutch:sc.log1p.Pref")) %>%
  mutate(Foodweb = c(rep("Complex",6),rep("Simple",6)),
         Selection_form = c(rep("Nonlinear",3),rep("Correlational",3),rep("Nonlinear",3),rep("Correlational",3)),
         Multiplier = c(rep(2,3),rep(1,3),rep(2,3),rep(1,3)))

# Combine and tidy the regression (alpha) coefficients
get_alphas <- bind_rows(foodweb_tidy_betas, foodweb_tidy_gammas) %>% 
  mutate(alpha = statistic, alpha_2.5 = conf.low, alpha_97.5 = conf.high, P = ifelse(conf.low*conf.high > 0, "<0.05","NS")) %>% # change 'alpha = estimate' to 'alpha = statistic' for bootstrap. Multiply confidence intervals to determine whether the cofficient is signficantly different from zero.
  select(term, Foodweb, alpha, alpha_2.5, alpha_97.5, P, Selection_form, Multiplier)

@@
